Multi-target detection with application to cryo-electron microscopy

### 2019-03-05 inverse_problems/inverse_problems.tex
sec 3.2

In this section, we give the expressions for the autocorrelations of the observed signal $y$ under the Poisson model. %In this model, each location $k$, $1 \le k \le N-L+1$ is the starting point of a Poisson-distributed number of random signals. 
%More precisely, letting $s[k]$ be the number of signals starting at location $k$, $s[k] \overset{iid}{\sim} \Poisson(\gamma / L)$.
 We first note that the expected total number of signals occurring in $y$ is equal to $(N - L + 1)\gamma / L$, and consequently the total lengths of all signals (including overlaps) divided by $N$ is equal to
%
\begin{align}
\frac{N - L + 1}{N} \cdot L \cdot \frac{\gamma  }{L} 
\quad \overset{N\to\infty}{\longrightarrow} \quad \gamma.
\end{align}
%
Therefore, in the large $N$ limit $\gamma$ may be interpreted as the signal density, as in the well-separated model.

In Section~\ref{sec:proof_prop_poisson}, we prove the following expressions for the autocorrelations of $y$ under the Poisson model. As in the well-separated model, the observed autocorrelations do not depend on individual occurrences of the signal, but only on $x$ itself, and the parameters $\gamma$ and $\sigma$:
%
\begin{align}
a_y^1  &=  \gamma a_{x}^1, 
\label{eq:mean_micrograph2} \\
a_y^2[\ell] &=  (\gamma a_x^1)^2 + \gamma a_x^2[\ell] + \sigma^2 \ell[\ell], 
\label{eq:ac2_micrograph2} \\
a_y^3[\ell_1,\ell_2]  & = (\gamma a_x^1)^3 + \gamma a_x^1  
\cdot ( \gamma a_x^2[\ell_1] + \gamma a_x^2[\ell_2]
+ \gamma a_x^2[\ell_2-\ell_1]) + \gamma a_x^3[\ell_1,\ell_2] 
\nonumber \\
& \qquad + \gamma a_x^1 \sigma^2 \big(\ell[\ell_1]+\ell[\ell_2] +\ell[\ell_1-\ell_2] \big).
\label{eq:ac3_micrograph2}
\end{align}

@@

We will denote by $\M_l$ the moment tensors of $x$:
%
\begin{align}
%
\M_1[i] = \E x[i], \quad 0 \le i \le L-1,
%
\end{align}
%
\begin{align}
%
\M_2[i,j] = \E x[i] x[j], \quad 0 \le i,j \le L-1,
%
\end{align}
%
\begin{align}
%
\M_3[i,j,k] = \E x[i] x[j] x[k], \quad 0 \le i,j,k \le L-1.
%
\end{align}

We obtain the autocorrelations $a_x^l$ of $x$ by averaging over a slice of the moment tensors:
%
\begin{align}
%
a_x^1 = \frac{1}{L}\sum_{i=0}^{L-1} \M_1[i],
%
\end{align}
%
\begin{align}
%
a_x^2[\ell] = \frac{1}{L} \sum_{i=0}^{L-1} \M_2[i,i+\ell],
%
\end{align}
%
and \TODO{Compare with~\eqref{eq:ac_special}: the limits on the sum should be modified}
%
\begin{align}
%
a_x^3[\ell_1,\ell_2] = \frac{1}{L}\sum_{i=0}^{L-1} \M_3[i,i+\ell_1,i+\ell_2].
%
\end{align}

@@

First we will consider the noise-free case, where $\sigma = 0$. We will condition on $s$ first, and then take the expectation over $s$. Fix $i_1 \ne i_2$, and let $\ell = i_2 - i_1$. Then:
%
\begin{align}
%
y[i_1] y[i_2]
&= \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
\sum_{k_1=1}^{s[i_1-j_1]}\sum_{k_2=1}^{s[i_2-j_2]}
x_{(k_1)}^{i_1-j_1}[j_1] x_{(k_2)}^{i_2 - j_2}[j_2].
%
\end{align}

@@

For term (b), the expectation conditional on $S$ is:
%
\begin{align}
%
\sum_{j=0}^{L-1} s[q] (s[q] - 1) \M_2[j,j+\ell_1] \M_1[j + \ell_2]
%
\end{align}
%
and then again using Lemma~\ref{lem-choose} we get the expected value:
%
\begin{align} \label{bbbb}
%
\frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \M_2[j,j+\ell_1] \M_1[j+\ell_2].
%
\end{align}

Similarly, the expected values of terms (c) and (d) are:
%
\begin{align} \label{cccc}
%
\frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \M_2[j,j+\ell_2] \M_1[j+\ell_1].
%
\end{align}
%
and
%
\begin{align} \label{dddd}
%
\frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \M_2[j+\ell_1,j+\ell_2] \M_1[j].
%
\end{align}

@@

Moving onto Case~\ref{case2}, we have $q_1 \equiv i_1 - j_1 = i_2 - j_2$, and also define $q_2 \equiv i_3 - j_3$. By definition, $q_1 \ne q_2$. The sum is:
%
\begin{align}
%
& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \ell_2}
\sum_{1 \le k_1,k_2 \le S[q_1]} \sum_{k_3=1}^{s[q_2]}
x_{(k_1)}^{q_1}[j_1] x_{(k_2)}^{q_1}[j_1 + \ell_1] x_{(k_3)}^{q_2}[j_3]   
\nonumber \\
=& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \ell_2} \sum_{k_3=1}^{s[q_2]}
\Bigg\{  \sum_{1 \le k_1 \ne k_2 \le s[q_1]} 
x_{(k_1)}^{q_1}[j_1] x_{(k_2)}^{q_1}[j_1 + \ell_1] x_{(k_3)}^{q_2}[j_3]
\nonumber \\
& + \sum_{k_1=1}^{s[q_1]} x_{(k_1)}^{q_1}[j_1] x_{(k_1)}^{q_1}[j_1 + \ell_1] 
x_{(k_3)}^{q_2}[j_3]  \Bigg\}.
%
\end{align}
%
Taking expectations conditional on $s$, we then get:
%
\begin{align}
%
\sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \ell_2} 
\Bigg( & s[q_1] (s[q_1]-1) s[q_2] \M_1[j_1] \M_1[j_1 + \ell_1] \M_1[j_3]
 + s[q_1] s[q_2] \M_2[j_1,j_1+\ell_1] \M_1[j_3] \Bigg).
%
\end{align}

>>>>

In this section, we give the expressions for the autocorrelations of the observed signal $y$ under the Poisson model. %In this model, each location $k$, $1 \le k \le N-L+1$ is the starting point of a Poisson-distributed number of random signals. 
%More precisely, letting $s[k]$ be the number of signals starting at location $k$, $s[k] \overset{iid}{\sim} \Poisson(\gamma / L)$.
 We first note that the expected total number of signals occurring in $y$ is equal to $(N - L + 1)\gamma / L$, and consequently the total lengths of all signals (including overlaps) divided by $N$ is equal to
%
\begin{align}
\frac{N - L + 1}{N} \cdot L \cdot \frac{\gamma  }{L} 
\quad \overset{N\to\infty}{\longrightarrow} \quad \gamma.
\end{align}
%
Therefore, in the large $N$ limit $\gamma$ may be interpreted as the signal density, as in the well-separated model~\eqref{eq:gamma}.

In Section~\ref{sec:proof_prop_poisson}, we prove the following expressions for the autocorrelations of $y$ under the Poisson model. As in the well-separated model, the observed autocorrelations do not depend on individual occurrences of the signal, but only on $x$ itself, and the parameters $\gamma$ and $\sigma$:
%
\begin{align}
a_y^1  &=  \gamma a_{x}^1, 
\label{eq:mean_micrograph2} \\
a_y^2[\ell] &=  (\gamma a_x^1)^2 + \gamma a_x^2[\ell] + \sigma^2 \delta[\ell], 
\label{eq:ac2_micrograph2} \\
a_y^3[\ell_1,\ell_2]  & = (\gamma a_x^1)^3 + \gamma a_x^1  
\cdot ( \gamma a_x^2[\ell_1] + \gamma a_x^2[\ell_2]
+ \gamma a_x^2[\ell_2-\ell_1]) + \gamma a_x^3[\ell_1,\ell_2] 
\nonumber \\
& \qquad + \gamma a_x^1 \sigma^2 \big(\delta[\ell_1]+\delta[\ell_2] +\delta[\ell_1-\ell_2] \big).
\label{eq:ac3_micrograph2}
\end{align}

@@

We will denote by $\M_q$ the moment tensors of $x$:
%
\begin{align}
%
\M_1[i] = \E x[i], \quad 0 \le i \le L-1,
%
\end{align}
%
\begin{align}
%
\M_2[i,j] = \E x[i] x[j], \quad 0 \le i,j \le L-1,
%
\end{align}
%
\begin{align}
%
\M_3[i,j,k] = \E x[i] x[j] x[k], \quad 0 \le i,j,k \le L-1.
%
\end{align}
We obtain the autocorrelations $a_x^q$ of $x$ by averaging over a slice of the moment tensors:
%
\begin{align}
%
a_x^1 = \frac{1}{L}\sum_{i=0}^{L-1} \M_1[i],
%
\end{align}
%
\begin{align}
%
a_x^2[\ell] = \frac{1}{L} \sum_{i=0}^{L-1} \M_2[i,i+\ell],
%
\end{align}
%
and \TODO{Compare with~\eqref{eq:ac_special}: the limits on the sum should be modified}
%
\begin{align}
%
a_x^3[\ell_1,\ell_2] = \frac{1}{L}\sum_{i=0}^{L-1} \M_3[i,i+\ell_1,i+\ell_2].
%
\end{align}

@@

First we will consider the noise-free case, where $\sigma = 0$. 
Again, we will condition on $s$ first, and then take the expectation over $s$. Fix $i_1 \ne i_2$, and let $\ell = i_2 - i_1$. Then:
%
\begin{align}
%
y[i_1] y[i_2]
&= \sum_{j_1=0}^{L-1} \sum_{j_2=0}^{L-1} 
\sum_{k_1=1}^{s[i_1-j_1]}\sum_{k_2=1}^{s[i_2-j_2]}
x_{(k_1)}^{i_1-j_1}[j_1] x_{(k_2)}^{i_2 - j_2}[j_2].
%
\end{align}

@@

For term (b), the expectation conditional on $S$ is:
%
\begin{align}
%
\sum_{j=0}^{L-1} s[q] (s[q] - 1) \M_2[j,j+\ell_1] \M_1[j + \ell_2]
%
\end{align}
%
and then again using Lemma~\ref{lem-choose} we get the expected value:
%
\begin{align} \label{bbbb}
%
\frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \M_2[j,j+\ell_1] \M_1[j+\ell_2].
%
\end{align}
Similarly, the expected values of terms (c) and (d) are:
%
\begin{align} \label{cccc}
%
\frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \M_2[j,j+\ell_2] \M_1[j+\ell_1].
%
\end{align}
%
and
%
\begin{align} \label{dddd}
%
\frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \M_2[j+\ell_1,j+\ell_2] \M_1[j].
%
\end{align}

@@

Moving onto Case~\ref{case2}, we have $q_1 \equiv i_1 - j_1 = i_2 - j_2$, and also define $q_2 \equiv i_3 - j_3$. By definition, $q_1 \ne q_2$. The sum is:
%
\begin{align}
%
& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \ell_2}
\sum_{1 \le k_1,k_2 \le s[q_1]} \sum_{k_3=1}^{s[q_2]}
x_{(k_1)}^{q_1}[j_1] x_{(k_2)}^{q_1}[j_1 + \ell_1] x_{(k_3)}^{q_2}[j_3]   
\nonumber \\
=& \sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \ell_2} \sum_{k_3=1}^{s[q_2]}
\Bigg\{  \sum_{1 \le k_1 \ne k_2 \le s[q_1]} 
x_{(k_1)}^{q_1}[j_1] x_{(k_2)}^{q_1}[j_1 + \ell_1] x_{(k_3)}^{q_2}[j_3]
\nonumber \\
& + \sum_{k_1=1}^{s[q_1]} x_{(k_1)}^{q_1}[j_1] x_{(k_1)}^{q_1}[j_1 + \ell_1] 
x_{(k_3)}^{q_2}[j_3]  \Bigg\}.
%
\end{align}
%
Taking expectations conditional on $s$, we then get:
%
\begin{align}
%
\sum_{j_1=0}^{L-1} \sum_{j_3 \ne j_1 + \ell_2} 
\Bigg( & s[q_1] (s[q_1]-1) s[q_2] \M_1[j_1] \M_1[j_1 + \ell_1] \M_1[j_3]
 + s[q_1] s[q_2] \M_2[j_1,j_1+\ell_1] \M_1[j_3] \Bigg).
%
\end{align}
